---
title: "AI Exposure Join"
subtitle: "Matching job postings to SOC codes and AI exposure scores"
author: "Richie"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
```

## Step 1: Load the crosswalk

This crosswalk maps keyword patterns and job titles to standardized SOC codes.

```{r}
#| label: load-crosswalk
crosswalk <- read_csv("job_soc_crosswalk.csv")
glimpse(crosswalk)
```

## Step 2: Load AI exposure data

Download from [Anthropic's HuggingFace repo](https://huggingface.co/datasets/Anthropic/EconomicIndex)
under the **"Labor market impacts"** folder. Rename the file to match below.

```{r}
#| label: load-exposure
# Option A: Anthropic labor market impacts (observed exposure scores by SOC)
anthropic_exposure <- read_csv("occupation_exposure.csv")  # rename to match actual file

# Option B: Eloundou et al. task-level beta scores (release_2025_02_10/ folder)
# eloundou_tasks <- read_csv("task_exposure_eloundou.csv")

glimpse(anthropic_exposure)
```

## Step 3: Load your job postings

Assumes a column called `job_title` exists in your data.

```{r}
#| label: load-jobs
your_jobs <- read_csv("your_job_postings.csv")  # replace with your file

glimpse(your_jobs)
```

## Step 4: Assign SOC codes via keyword matching

Multi-word specific patterns are matched first (e.g., `"data scientist"`), then
single-keyword fallbacks (e.g., `"data"`).

```{r}
#| label: assign-soc
crosswalk_primary <- crosswalk %>%
  filter(match_confidence %in% c("High", "Medium")) %>%
  arrange(soc_code, match_confidence) %>%
  group_by(keyword_pattern) %>%
  slice(1) %>%
  ungroup()

assign_soc <- function(title) {
  title_lower <- str_to_lower(title)

  # Try multi-word (specific) patterns first
  specific_patterns <- crosswalk_primary %>%
    filter(str_detect(keyword_pattern, "\\s"))

  for (i in seq_len(nrow(specific_patterns))) {
    if (str_detect(title_lower, specific_patterns$regex_example[i])) {
      return(specific_patterns$soc_code[i])
    }
  }

  # Fall back to single-keyword patterns
  keyword_patterns <- crosswalk_primary %>%
    filter(!str_detect(keyword_pattern, "\\s"))

  for (i in seq_len(nrow(keyword_patterns))) {
    if (str_detect(title_lower, keyword_patterns$regex_example[i])) {
      return(keyword_patterns$soc_code[i])
    }
  }

  return(NA_character_)
}

your_jobs <- your_jobs %>%
  mutate(soc_code = map_chr(job_title, assign_soc))
```

## Step 5: Join to AI exposure scores

```{r}
#| label: join-exposure
jobs_with_exposure <- your_jobs %>%
  left_join(anthropic_exposure, by = "soc_code")
```

### Coverage check

How many of your job postings successfully matched to an exposure score?

```{r}
#| label: coverage
jobs_with_exposure %>%
  summarise(
    total_jobs       = n(),
    matched_soc      = sum(!is.na(soc_code)),
    matched_exposure = sum(!is.na(observed_exposure)),  # update col name to match actual file
    pct_matched      = round(matched_exposure / total_jobs, 3)
  )
```

## Step 6: (Optional) All matching SOC codes per job title

Useful if you want to flag that a job could reasonably map to multiple occupations.

```{r}
#| label: multi-soc
jobs_multi_soc <- your_jobs %>%
  select(job_title) %>%
  distinct() %>%
  mutate(
    matched_socs = map(job_title, function(title) {
      title_lower <- str_to_lower(title)
      crosswalk %>%
        filter(map_lgl(regex_example, ~ str_detect(title_lower, .x))) %>%
        select(soc_code, soc_title, match_confidence)
    })
  ) %>%
  unnest(matched_socs)

jobs_multi_soc
```

## Step 7: Summary table — mean exposure by SOC code

```{r}
#| label: summary
exposure_summary <- jobs_with_exposure %>%
  group_by(soc_code) %>%
  summarise(
    n_jobs        = n(),
    mean_exposure = mean(observed_exposure, na.rm = TRUE),  # update col name
    .groups = "drop"
  ) %>%
  left_join(
    crosswalk %>% select(soc_code, soc_title) %>% distinct(),
    by = "soc_code"
  ) %>%
  arrange(desc(mean_exposure))

exposure_summary
```
